### 2019-09-17 reports/thesis/chapters/3_compoundPoisson.tex
Revise

Returning back to the compound Poisson with no electronic noise $X|Y = \sum_{i=1}^{Y}U_i$, let the moment generating function (m.g.f.) of $X$ be $M_X(\theta)=\expectation\left[\euler^{X\theta}\right]$. It can be shown that the m.g.f.~is
\begin{equation}
  M_X(\theta)=
  \exp\left[
    \lambda
    \left(
      M_U(\theta)-1
    \right)
  \right]
\end{equation}
\citep{gatto2010saddlepoint}. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}. Moments of $X$ can be obtained from the m.g.f.~by differentiating it and setting $\theta$ to zero. In other words, $\expectation[X^r]=M_X^{(r)}(0)$. Then it can be shown that
\begin{align}
  \expectation\left[X\right]&=\lambda\expectation\left[U\right]
  \\
  \variance\left[X\right] &= \lambda\expectation\left[U^2\right]
  \\
  \expectation\left[(X-\expectation[X])^3\right] &=\lambda\expectation\left[U^3\right]
\end{align}
which agrees with the expectation and variance results in the previous section. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}.

\subsection{Compound Poisson-Gamma Distribution}
\label{chapter:compoundPoisson_compoundPoissonGamma}
A special case of the compound Poisson is when $U\sim\gammaDist\left(\alpha,\beta\right)$ where $\alpha>0$ is the gamma shape parameter and $\beta>0$ is the gamma rate parameter. This was used for example in \cite{xu2009electronic}. The distribution is known as the compound Poisson-gamma distribution and is denoted by $X\sim\CPoisson(\lambda,\alpha,\beta)$. The p.d.f.~is
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda} & \text{ for } x=0 \\ 
    \displaystyle\sum_{y=1}^{\infty}\dfrac{\beta^{y\alpha}}{\Gamma(y\alpha)}x^{y\alpha-1}\euler^{-\beta x}\euler^{-\lambda}\frac{\lambda^y}{y!} & \text{ for } x>0
  \end{cases}
  \ .
  \label{eq:compoundPoisson_pdf}
\end{equation}
Recall that $X|Y=\sum_{i=1}^YU_i$ which involves a sum of gamma random variables. It can be shown that
\begin{equation}
  X|Y\sim\gammaDist\left(Y\alpha,\beta\right).
\end{equation}

@@

Let
\begin{equation}
  W_y^{(r)} = y^r W_y \quad \text{for }r=1,2,3,\dotsc
\end{equation}
so that
\begin{equation}
  y_i = \frac{\sum_{y=1}^\infty W^{(1)}_y}{\sum_{y=1}^\infty W_y} \ .
\end{equation}
Similarly
\begin{equation}
  \zeta_i = \variance[Y|X=x_i] = \frac{\sum_{y=1}^\infty W^{(2)}_y}{\sum_{y=1}^\infty W_y} - \left(y_i\right)^2
  \ .
\end{equation}
The expectation terms, $y_i$ and $\zeta_i$, are evaluated in the M step so that they can be used in the E step. The evaluation can be done by truncating the sum
\begin{equation}
  \sum_{y=1}^\infty W^{(r)}_y \approx \sum_{y=y_\text{l}}^{y_\text{u}} W^{(r)}_y
\end{equation}
where $y_\text{l}<y_\text{max}<y_\text{u}$ and $y_\text{max}$ is the value of $y$ which maximises $W_y^{(r)}$. $\ln(W^{(r)}_y)$ can be expressed as
\begin{equation}
  \ln\left(W_y^{(r)}\right)=r\ln(y)+\ln(W_y)
\end{equation}
and taking the derivative with respect to $y$ obtains
\begin{equation}
  \frac{\partial \ln(W_y^{(r)})}{\partial y} = \frac{r}{y } + \frac{\partial \ln(W_y)}{\partial y}
  \ .
\end{equation}
Keep in mind that $y=1,2,3,\dotsc$ so for large $y$, an approximation can be made $r/y\approx 0$ so that
\begin{equation}
  \frac{\partial \ln(W_y^{(r)})}{\partial y} \approx \frac{\partial \ln(W_y)}{\partial y}
  \ .
\end{equation}
Therefore, the maximum of $W_y^{(r)}$ is located at $y_\text{max}$ for all $r=0,1,2,\dotsc$. As a results, the same method for evaluating $W_y$ can be used to evaluate $W_y^{(r)}$. For a given $r=0,1,2,\dotsc$, the limit of the sum $\sum_{y=y_\text{l}}^{y_\text{u}} W^{(r)}_y$ are chosen such that $W_{y_\text{l}}$ and $W_{y_\text{u}}$ are less than $\epsilon W_{y_\text{max}}^{(r)}$ where $\epsilon$ is some small constant. The limits will be different for different values of $r$.

@@

In the M step, the conditional expected joint log-likelihood is maximised with respect to the parameters $\lambda$, $\alpha$ and $\beta$. The objective function is
\begin{align}
  T(\lambda,\alpha,\beta)&=
  \begin{multlined}[t]
    \sum_{i=1}^n
    \expectation\left[
      -\mathbb{I}(x_i=0)
      \lambda
    \right.
    \\
    \left.+
      \mathbb{I}(x_i>0)
      \left(
        Y_i\alpha\ln\beta-\ln\Gamma(Y_i\alpha)+(Y_i\alpha-1)\ln x_i - \beta x_i
      \right.
    \right.
    \\
    \left.
      \left.  
        - \lambda + Y_i \ln \lambda - \ln(Y_i!)
      \right)
      |X_i=x_i
    \right]
  \end{multlined}
  \nonumber\\
  &\!\begin{multlined}[b]=
    -n\lambda
    \\
    +\sum_{i=1}^n
    \mathbb{I}(x_i>0)
    \left[
      \expectation[Y_i|X_i=x_i]\alpha\ln\beta-\expectation[\ln\Gamma(Y_i\alpha)|X_i=x_i]
    \right.
    {}\\
    \left.
      +\expectation[Y_i|X_i=x_i]\alpha\ln x_i - \beta x_i
      + \expectation[Y_i|X_i=x_i] \ln \lambda
    \right] + c
  \end{multlined}
\end{align}
where $c$ is some constant not dependent on $\lambda$, $\alpha$ or $\beta$.

>>>>

Returning back to the compound Poisson distribution with no electronic noise $X|Y = \sum_{i=1}^{Y}U_i$, let the moment generating function (m.g.f.) of $X$ be $M_X(\theta)=\expectation\left[\euler^{X\theta}\right]$. It can be shown that the m.g.f.~is
\begin{equation}
  M_X(\theta)=
  \exp\left[
    \lambda
    \left(
      M_U(\theta)-1
    \right)
  \right]
\end{equation}
\citep{gatto2010saddlepoint}. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}. Moments of $X$ can be obtained from the m.g.f.~by differentiating it and setting $\theta$ to zero. In other words, $\expectation[X^r]=M_X^{(r)}(0)$. Then it can be shown that
\begin{align}
  \expectation\left[X\right]&=\lambda\expectation\left[U\right]
  \\
  \variance\left[X\right] &= \lambda\expectation\left[U^2\right]
  \\
  \expectation\left[(X-\expectation[X])^3\right] &=\lambda\expectation\left[U^3\right]
\end{align}
which agrees with the expectation and variance results in the previous section. The derivation is shown in Appendix \ref{chapter:appendix_compoundPoissonMgf}.

\subsection{Compound Poisson-Gamma Distribution}
\label{chapter:compoundPoisson_compoundPoissonGamma}
A special case of the compound Poisson distribution is when $U\sim\gammaDist\left(\alpha,\beta\right)$ where $\alpha>0$ is the gamma shape parameter and $\beta>0$ is the gamma rate parameter. This was used for example in \cite{xu2009electronic}. The distribution is known as the compound Poisson-gamma distribution and is denoted by $X\sim\CPoisson(\lambda,\alpha,\beta)$. The p.d.f.~is
\begin{equation}
  p_X(x) = 
  \begin{cases}
    \delta(x) \euler^{-\lambda} & \text{ for } x=0 \\ 
    \displaystyle\sum_{y=1}^{\infty}\dfrac{\beta^{y\alpha}}{\Gamma(y\alpha)}x^{y\alpha-1}\euler^{-\beta x}\euler^{-\lambda}\frac{\lambda^y}{y!} & \text{ for } x>0
  \end{cases}
  \ .
  \label{eq:compoundPoisson_pdf}
\end{equation}
Recall that $X|Y=\sum_{i=1}^YU_i$ which involves a sum of gamma random variables. It can be shown that
\begin{equation}
  X|Y\sim\gammaDist\left(Y\alpha,\beta\right).
\end{equation}

@@

Let
\begin{equation}
  W_y^{(r)} = y^r W_y \quad \text{for }r=1,2,3,\dotsc
\end{equation}
so that
\begin{equation}
  y_i = \frac{\sum_{y=1}^\infty W^{(1)}_y}{\sum_{y=1}^\infty W_y} \ .
\end{equation}
Similarly,
\begin{equation}
  \zeta_i = \variance[Y|X=x_i] = \frac{\sum_{y=1}^\infty W^{(2)}_y}{\sum_{y=1}^\infty W_y} - \left(y_i\right)^2
  \ .
\end{equation}
The expectation terms, $y_i$ and $\zeta_i$, are evaluated in the M step so that they can be used in the E step. The evaluation can be done by truncating the sum
\begin{equation}
  \sum_{y=1}^\infty W^{(r)}_y \approx \sum_{y=y_\text{l}}^{y_\text{u}} W^{(r)}_y
\end{equation}
where $y_\text{l}<y_\text{max}<y_\text{u}$ and $y_\text{max}$ is the value of $y$ which maximises $W_y^{(r)}$. $\ln(W^{(r)}_y)$ can be expressed as
\begin{equation}
  \ln\left(W_y^{(r)}\right)=r\ln(y)+\ln(W_y)
\end{equation}
and taking the derivative with respect to $y$ obtains
\begin{equation}
  \frac{\partial \ln(W_y^{(r)})}{\partial y} = \frac{r}{y } + \frac{\partial \ln(W_y)}{\partial y}
  \ .
\end{equation}
Keep in mind that $y=1,2,3,\dotsc$ so for large $y$, an approximation can be made $r/y\approx 0$ so that
\begin{equation}
  \frac{\partial \ln(W_y^{(r)})}{\partial y} \approx \frac{\partial \ln(W_y)}{\partial y}
  \ .
\end{equation}
Therefore, the maximum of $W_y^{(r)}$ is located at $y_\text{max}=\dfrac{x^{2-p}}{\phi(2-p)}$ for all $r=0,1,2,\dotsc$. As a result, the same method for evaluating $W_y$ can be used to evaluate $W_y^{(r)}$. For a given $r=0,1,2,\dotsc$, the limit of the sum $\sum_{y=y_\text{l}}^{y_\text{u}} W^{(r)}_y$ are chosen such that $W_{y_\text{l}}$ and $W_{y_\text{u}}$ are less than $\epsilon W_{y_\text{max}}^{(r)}$ where $\epsilon$ is some small constant. The limits will be different for different values of $r$.

@@

In the M step, the conditional expected joint log-likelihood is maximised with respect to the parameters $\lambda$, $\alpha$ and $\beta$. The objective function is
\begin{align}
  T(\lambda,\alpha,\beta)&=
  \begin{multlined}[t]
    \sum_{i=1}^n
    \expectation\left[
      -\mathbb{I}(x_i=0)
      \lambda
    \right.
    \\
    \left.+
      \mathbb{I}(x_i>0)
      \left(
        Y_i\alpha\ln\beta-\ln\Gamma(Y_i\alpha)+(Y_i\alpha-1)\ln x_i - \beta x_i
      \right.
    \right.
    \\
    \left.
      \left.  
        - \lambda + Y_i \ln \lambda - \ln(Y_i!)
      \right)
      |X_i=x_i
    \right]
  \end{multlined}
  \nonumber\\
  &=\begin{multlined}[t]
    -n\lambda
    \\
    +\sum_{i=1}^n
    \mathbb{I}(x_i>0)
    \left[
      \expectation[Y_i|X_i=x_i]\alpha\ln\beta-\expectation[\ln\Gamma(Y_i\alpha)|X_i=x_i]
    \right.
    {}\\
    \left.
      +\expectation[Y_i|X_i=x_i]\alpha\ln x_i - \beta x_i
      + \expectation[Y_i|X_i=x_i] \ln \lambda
    \right] + c
  \end{multlined}
\end{align}
where $c$ is some constant not dependent on $\lambda$, $\alpha$ or $\beta$.
